---
title: "Data Visualization in R Worksheet Answers"
output:
  html_document:
    highlight: tango
    toc: yes
  pdf_document:
    highlight: tango
    toc: yes
date: "November 27, 2014"
author: "Leland Taylor"
---

# Data Visualization in R 

Worksheet for Data Visualization in R class. The data we use is from the *Anscombe* dataset.

http://vincentarelbundock.github.io/Rdatasets/doc/car/Anscombe.html

The Anscombe data frame has 51 rows and 4 columns. The observations are the U. S. states plus Washington, D. C. in 1970.


This data frame contains the following columns:

* education
    * Per-capita education expenditures, dollars.
* income
    * Per-capita income, dollars.
* young
    * Proportion under 18, per 1000.
* urban
    * Proportion urban, per 1000.

# load packages

```{r package, eval=T, echo=TRUE, results='markup', error=FALSE }
#install.packages("car") # download data

library(car)
library(ggplot2)
library(reshape2)
```

# process the data 

When running this on your own, you can load the pre-processed data using *read.csv*

```{r process, eval=T, echo=TRUE, results='markup', error=FALSE }
# generate the data for boxplots
dat_bx = melt(Anscombe)
write.csv(dat_bx, file = "Anscombe_summarize.csv", row.names=F)

# generate the data for regressions
dat_lm = melt(Anscombe, id.vars = c("education"))
write.csv(dat_lm, file = "Anscombe_regression.csv", row.names=F)

# read in the data
dat_bx = read.csv("Anscombe_summarize.csv", header=T)
dat_lm = read.csv("Anscombe_regression.csv", header=T)
```

# challenge 1

Potential solutions to challenge 1

```{r c1, eval=T, echo=TRUE, results='markup', error=FALSE }
# boxplot
ggplot(dat_bx, aes(x=variable, y=value, fill=variable)) + geom_boxplot(outlier.shape = NA) + geom_jitter()

# regressions
ggplot(dat_lm, aes(x=value, y=education, color=variable)) + geom_point(aes(color=variable)) + stat_smooth(method="lm", color="black")

ggplot(dat_lm, aes(x=value, y=education, color=variable)) + geom_point() + stat_smooth(method="lm", se=F)
```


# challenge 2

Potential solutions to challenge 2

```{r c2, eval=T, echo=TRUE, results='markup', error=FALSE }
# boxplot
ggplot(dat_bx, aes(x=variable, y=value, fill=variable)) + geom_boxplot(outlier.shape = NA) + geom_jitter() + labs(x=NULL, y="Values", title="Distributions of data") + guides(fill=guide_legend(title=NULL))

ggplot(dat_bx, aes(x=variable, y=value, fill=variable)) + geom_boxplot(outlier.shape = NA) + geom_jitter() + labs(x=NULL, y="log10(Values)", title="Distributions of data") + guides(fill=guide_legend(title=NULL)) + scale_y_log10()
```

# challenge 3

Potential solutions to challenge 3

```{r c3, eval=T, echo=TRUE, results='markup', error=FALSE }
# initialize the plot
plt = ggplot(dat_lm, aes(x=value, y=education, color=variable)) + geom_point() 

# add a linear regression
plt = plt + stat_smooth(method="lm") 

# set axis labels
plt = plt + labs(x="Value", 
                 y="Per-capita education expenditures", 
                 title="Linear regression (edu ~ value)")

# set guide label
plt = plt + guides(color=guide_legend(title="Variable")) 

# change the theme
plt = plt + theme_bw() 

# EBI colors :)
plt = plt + scale_color_manual(values=c("#71B360", "#48877C", "#CC6D78")) 

# wrap
plt = plt + facet_wrap(~ variable, scales="free") 

# Adjust the x axis so we can read them
plt = plt + theme( axis.text.x=element_text( angle=-45, hjust=0 ) ) 

# show the plot to the screen
show(plt)


# save the plot to a pdf
pdf(file="challenge_3.pdf", height=5, width=6)
  print(plt)
dev.off()
```

